\documentclass[tikz,border=8pt]{standalone}

\usepackage{fontspec}
\usepackage{xeCJK}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,backgrounds,calc,fit,positioning}

\setmainfont{Linux Libertine O}
\setsansfont{TeX Gyre Heros}
\setmonofont{Latin Modern Mono}
\setCJKmainfont{Noto Serif CJK TC}
\setCJKsansfont{Noto Serif CJK TC}
\renewcommand{\familydefault}{\rmdefault}
\hyphenpenalty=10000
\exhyphenpenalty=10000
\pagestyle{empty}

\newcommand{\SharedLaneTitle}{Train/valid Setup}
\newcommand{\TrainingLaneTitle}{Training Only}
\newcommand{\InferenceLaneTitle}{Test-Time Prediction}
\newcommand{\LegendBoxTitle}{Legend}
\newcommand{\LegendDataLabel}{Data}
\newcommand{\LegendProcessLabel}{Process}
\newcommand{\LegendModelLabel}{Model}
\newcommand{\LegendOutputLabel}{Output/post-process}
\newcommand{\LegendMainFlowLabel}{Main flow}
\newcommand{\LegendReuseLabel}{Reused dependency}
\newcommand{\LegendReuseNote}{gray text = reused}
\newcommand{\ReusePrefix}{Reuse: }
\newcommand{\EpochLoopLabel}{repeat each epoch}

\newcommand{\LegendTitle}{How to read this figure}
\newcommand{\LegendBody}{Solid arrows are the main flow.\\Gray text marks reused results from earlier steps.}

\newcommand{\RawDataTitle}{Raw train cases\\and labels}
\newcommand{\RawDataSub}{Input: training case files and noticed-case labels.}

\newcommand{\PreprocessTitle}{Clean the text}
\newcommand{\PreprocessSub}{Output: cleaned text.}

\newcommand{\SplitAssetsTitle}{Split train/valid\\labels}
\newcommand{\SplitAssetsSub}{Output: split labels and valid query IDs.}

\newcommand{\IndexScopeTitle}{Build train-valid\\BM25 and scope}
\newcommand{\IndexScopeSub}{Output: train\_qid, BM25 query/index/ranking, and candidate scope.}

\newcommand{\BackboneTitle}{ModernBERT\\backbone}
\newcommand{\BackboneSub}{Start from the base encoder or a legal DAPT checkpoint.}

\newcommand{\HardNegTitle}{Prepare contrastive\\data}
\newcommand{\HardNegSub}{Load train positives and a fixed validation set.}

\newcommand{\DynamicNegTitle}{Sample train\\negatives}
\newcommand{\DynamicNegSub}{At each epoch, resample 15 train negatives from current dense scores.}

\newcommand{\SFTTitle}{Contrastive SFT\\and dense encoder}
\newcommand{\SFTSub}{Input: query, positive, and 15 negatives.\\Output: trained dense encoder.}

\newcommand{\LTRTrainTitle}{Train the LTR\\reranker}
\newcommand{\LTRTrainSub}{Build 28 features per pair and train LightGBM LambdaRank on train-valid.}

\newcommand{\InferenceInputTitle}{Prepare test\\assets}
\newcommand{\InferenceInputSub}{Clean test text and build test qids, BM25 inputs/index, and year scope.}

\newcommand{\LexicalTitle}{Lexical scoring}
\newcommand{\LexicalSub}{Compute BM25 and QLD scores within the test scope.\ReuseHint{test BM25 / scope}}

\newcommand{\DenseTitle}{Dense scoring}
\newcommand{\DenseSub}{Encode query-candidate texts into embeddings, then score dense similarity.\ReuseHint{trained encoder}}

\newcommand{\FeatureMergeTitle}{LTR reranking}
\newcommand{\FeatureMergeSub}{Build the same 28 features from lexical and dense signals, then rerank.\ReuseHint{best LTR model}}

\newcommand{\RerankTitle}{Best LightGBM\\reranker}
\newcommand{\RerankSub}{Use the trained LambdaRank model to rerank each query's candidates.}

\newcommand{\PostTitle}{Post-process\\and export}
\newcommand{\PostSub}{Remove self-cases, choose cutoff, and export the final answer.}

\input{task1_system_pipeline_common.tex}

\begin{document}
\TaskOneSystemPipelineFigure
\end{document}
