\section{Introduction}

Legal case retrieval is a core task in legal information systems. Given a new case document as query, the system must retrieve the most relevant prior cases to support legal research, citation analysis, and judicial decision making. COLIEE Task~1 has long provided a standardized benchmark for this setting, enabling controlled comparison across retrieval methods.

Our study starts from the observation reported by THUIR@COLIEE 2023 \cite{li2023thuir}. Their Table~3 shows that SAILER outperforms BM25 when both are constrained to 512 tokens, but unrestricted BM25 becomes stronger again than truncated SAILER. This suggests that the relative performance gap in legal retrieval may depend not only on model architecture, but also on how much of the case document the model is actually allowed to read.

COLIEE 2026 itself is strongly long-document oriented. As shown in Table~\ref{tab:length-stats}, the 7,708 \texttt{processed} cases have an average length of 5,031.36 tokens and a median of 3,684.5 tokens; 56.11\% are shorter than 4096 tokens, 87.69\% are shorter than 8192 tokens, and 94.80\% are shorter than 12,288 tokens. This directly motivates our later chunk-based reranking design: if each document is split into at most three 4096-token chunks, almost the entire corpus can still be covered without increasing the encoder limit beyond 4096.

\begin{table}[t]
  \centering
  \small
  \begin{tabular}{lr}
    \toprule
    Statistic & Value \\
    \midrule
    Documents & 7,708 \\
    Mean tokens & 5,031.36 \\
    Median tokens & 3,684.5 \\
    Max tokens & 144,941 \\
    $<4096$ tokens & 4,325 (56.11\%) \\
    $<8192$ tokens & 6,759 (87.69\%) \\
    $<12,288$ tokens & 7,307 (94.80\%) \\
    \bottomrule
  \end{tabular}
  \caption{Length statistics of COLIEE 2026 \texttt{processed} documents. Since our learning-to-rank stage uses at most three 4096-token chunks, 12,288 tokens is the effective coverage limit for chunk-based reranking.}
  \label{tab:length-stats}
\end{table}

To make the truncation issue more concrete, Table~\ref{tab:thuir-table3-excerpt} reproduces key numbers from THUIR@COLIEE 2023 Table~3. When the maximum length is fixed at 512 tokens, SAILER achieves F1 of 0.1385 and BM25 only 0.1012. However, when BM25 is not length-limited, its F1 rises to 0.1541 and becomes stronger than truncated SAILER. These numbers alone do not prove that dense retrieval will always win given longer context, but they do strongly support our central hypothesis: long-text truncation is one of the main reasons why dense retrieval cannot stably surpass lexical retrieval in legal case retrieval.

\begin{table}[t]
  \centering
  \small
  \begin{tabular}{lcccc}
    \toprule
    Model & max\_length & P@5 & R@5 & F1 \\
    \midrule
    BM25($k_1=3,b=1$) & 512 & 0.0963 & 0.1067 & 0.1012 \\
    SAILER & 512 & 0.1315 & 0.1459 & 0.1385 \\
    BM25($k_1=3,b=1$) & -- & 0.1465 & 0.1625 & 0.1541 \\
    \bottomrule
  \end{tabular}
  \caption{Key rows excerpted from THUIR@COLIEE 2023 Table~3 \cite{li2023thuir}. ``--'' indicates unlimited length.}
  \label{tab:thuir-table3-excerpt}
\end{table}

Based on this hypothesis, we choose ModernBERT \cite{warner2024modernbert} as our long-context encoder and further apply domain-adaptive pretraining on U.S. case law. In practice, however, hardware immediately becomes the next bottleneck. Under contrastive training with \texttt{query:positive:negatives=1:1:15}, using 8192 tokens per document causes CUDA out-of-memory on a single NVIDIA RTX 4080 SUPER 16GB, so all formal experiments use 4096 tokens instead. The masked language modeling stage over U.S. cases also requires about 60 hours while still covering only 0.384 epoch. Our work is therefore both a method study and an engineering study under realistic single-GPU constraints.

During development, we found that simply switching to a long-context encoder is not enough. THUIR-style reference sentence extraction degrades retrieval in our setting; ModernBERT without supervised fine-tuning is almost unusable; and a static BM25 hard-negative design similar to the THUIR pipeline fails to converge. Only after we replace static negatives with dynamic negative sampling, and then integrate year-based scope filtering, learning to rank, and post-processing, does the system become stable.

This paper makes three main contributions. First, we re-evaluate the long-context hypothesis on COLIEE 2026 and show that truncation length interacts strongly with query representation. Second, we demonstrate that static BM25 hard negatives can make legal dense retrieval nearly impossible to train, whereas dynamic negative sampling is the key to successful convergence. Third, we integrate dense retrieval, scope filtering, LightGBM learning to rank, and cut-off based post-processing into a complete system, which finally ranks 7th in the official COLIEE 2026 Task~1 results \cite{coliee2026task1results}.
