\section{Task and Data}

We use COLIEE 2026 Task~1 as the main benchmark and COLIEE 2025 as auxiliary data for methodological exploration and failure analysis. The goal of Task~1 is not to retrieve generally similar judgments, but to find the noticed cases for each query case. If a case was referenced by the query case, we treat it as a noticed case for that query. Because the organizers redact citation information from all case texts, the system must infer from the facts, reasoning, and legal semantics which prior cases most likely support the judgment.

\subsection{Data Sources and Splits}

The data come from the Federal Court of Canada. In the training data, the organizers provide query cases, the full candidate pool, and the gold noticed cases for each query. In the test data, only query cases are released, and the system must automatically predict the noticed cases. In other words, the whole pipeline from retrieval to final output must be fully automatic.

Table~\ref{tab:dataset-stats} summarizes the query counts, year-specific candidate pool sizes, and average numbers of noticed cases per query. COLIEE 2025 provides train and validation labels, whereas COLIEE 2026 additionally provides test labels. In our current setup, the COLIEE 2026 case collection contains 7,708 documents. We further split the official labels into 1,600 training queries and 401 validation queries. Model training and hyperparameter selection are performed on this train/validation split. Our main analysis reports top-5 micro-averaged F1, precision, and recall on both the 2026 validation and 2026 training queries. COLIEE 2025 is mainly used to document training failure, the effect of dynamic negatives, and the early gain from scope filtering.

\begin{table*}[t]
\caption{Dataset statistics of COLIEE Task~1. Candidate counts are measured from the year-specific \texttt{processed} pool.}
\label{tab:dataset-stats}
\centering
\small
\setlength{\tabcolsep}{5pt}
\begin{tabular}{lccccc}
\toprule
 & \multicolumn{2}{c}{COLIEE 2025} & \multicolumn{3}{c}{COLIEE 2026} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-6}
Statistic & Train & Valid & Train & Valid & Test \\
\midrule
\# of query cases & 1,341 & 336 & 1,600 & 401 & 400 \\
\# of candidate cases & 7,349 & 7,349 & 7,708 & 7,708 & 7,708 \\
Avg. \# of noticed cases per query & 4.07 & 4.24 & 4.12 & 4.13 & 4.38 \\
\bottomrule
\end{tabular}
\end{table*}

\subsection{\texttt{processed} and \texttt{processed\_new}}

We compare two query representations. The first is \texttt{processed}, which uses the full preprocessed case text as both query and candidate. The second is \texttt{processed\_new}, which follows THUIR's reference sentence extraction idea and keeps only local context around the citation-redaction placeholders; if the extracted text is too short, the script prepends the beginning of the full processed document. In our study, this is a query hypothesis that must be re-tested rather than accepted by default. For fair ablation, candidate documents are always encoded from \texttt{processed}; only the query representation changes.

\texttt{processed} is produced by \texttt{process.py}. It keeps the judgment body while removing surface forms of suppression placeholders such as \texttt{FRAGMENT\_SUPPRESSED}, editorial markers such as \texttt{[translation]}, \texttt{[sic]}, and \texttt{[End of document]}, and passages detected as French. If the cleaned body still lacks \texttt{Summary:}, \texttt{process.py} prepends a same-name summary file only when \texttt{summary.py} has already extracted that summary from the raw case. In other words, \texttt{summary.py} does not generate new summaries; it only copies an existing \texttt{Summary:} block from the raw file when such a block is present. \texttt{processed\_new} is produced by \texttt{reference.py}. It first extracts the sentences around citation-redaction placeholders and then applies nearly the same cleanup steps; if the result is still shorter than 512 English word tokens, it is expanded with the beginning of the \texttt{processed} version. Therefore, \texttt{processed\_new} is not a pure citation fragment but a reference-centered query view with fallback expansion from the full cleaned document.

\subsection{Long-Document Characteristics and Continued Pretraining Data}

As shown in Table~\ref{tab:length-stats}, COLIEE 2026 \texttt{processed} documents are clearly long, and 94.80\% of them are shorter than 12,288 tokens. This makes three 4096-token chunks a practical reranking approximation for most cases.

Besides the official COLIEE data, we use the Common Pile CaseLaw Access Project subset from Hugging Face\footnote{\url{https://huggingface.co/datasets/common-pile/caselaw_access_project_filtered}} for continued pretraining. It contains 6,735,525 U.S. case documents spanning roughly 365 years. This corpus does not provide direct COLIEE supervision; instead, it helps ModernBERT adapt to legal text distributions before contrastive fine-tuning. Under our current hardware budget, masked language modeling on this corpus runs for about 60 hours and still covers only 0.384 epoch, so we treat it as a useful but clearly under-trained adaptation stage.
