\section{Limitations and Future Work}

The first limitation is hardware. Although ModernBERT natively supports 8192 tokens, a contrastive setup with one query case, one noticed case, and 15 negatives cannot be trained stably at 8192 tokens on a single RTX 4080 SUPER 16GB, so the final system uses 4096 tokens. This means the present work still does not fully answer whether a truly 8192-token legal dense retriever would yield larger gains. Still, the document-length distribution suggests that it probably would: full-document coverage would rise from 56.11\% at 4096 tokens to 87.69\% at 8192 tokens, an additional 31.58 percentage points or 2,434 more judgments.

Second, the domain-adaptive pretraining stage on U.S. case law remains under-trained. After roughly 60 hours, the model has seen only 0.384 epoch of the 6,735,525-case corpus. Its potential is therefore unlikely to be exhausted. In addition, our DAP stage uses bf16, and we have not systematically compared fp16, bf16, fp32, or other mixed-precision settings for continued pretraining and downstream retrieval.

Third, year extraction still relies on a rule-based method. The current implementation scans each case for four-digit years from 1800 to 2026, takes the maximum matched year as the judgment year, and then builds query-specific scope with \texttt{candidate\_year <= query\_year + 1}. This rule is simple and effective, but it can still be affected by historical years, procedural dates, or preprocessing artifacts. If the organizers could release richer metadata such as decision date, jurisdiction, court level, judge, and procedural type, both candidate scope and other structured features could be defined more accurately.

Finally, the current reranking pipeline still relies mainly on a bi-encoder and a query-wise tree ranker. Future work can compare stronger ensemble methods and objective functions, and can also incorporate cross encoders, late interaction models such as ColBERT, and finer-grained paragraph-, sentence-, or token-level signals to test whether full-document noticed-case supervision can better exploit such detailed evidence.
