\section{Conclusion}

Starting from THUIR@COLIEE 2023, this paper re-examines long-document handling, query representation, negative design, and downstream ranking integration in legal case retrieval. We show that when the full judgment is used as the query, ModernBERT + DAP + SFT reaches 0.1650 F1@5 on COLIEE 2026 validation, outperforming both BM25 at 0.1499 and SAILER at 0.1213. In contrast, when the query is reduced to a reference-centered view, both neural models degrade substantially. This means that the advantage of a long-context encoder depends on actually preserving enough full-document context.

We also show that a long-context encoder is not a plug-and-play solution. ModernBERT without SFT is almost unusable for this task, and training with static BM25 hard negatives nearly collapses to random guessing. The real turning point is negative construction: we replace static hard negatives with epoch-wise dynamic negative sampling and also impose query-specific temporal scope constraints so that future cases are excluded from the negative pool. This lets the model learn genuinely un-noticed cases as negatives rather than merely legally impossible future cases.

Finally, dense retrieval is not the endpoint of the system. Once lexical, dense, length, placeholder, year, and chunk signals are passed to LightGBM learning to rank and combined with cut-off post-processing, FLNLP reaches 0.2935 F1 on COLIEE 2026 Task~1 and ranks 7th among 22 teams. The core challenge in legal dense retrieval is therefore not only choosing a larger model, but also preserving the right query representation, constructing informative negatives, and integrating retrieval signals into the final ranking stage under realistic hardware constraints.
